Flaring star-planet interactions in star-planet systems in Kepler and TESS

### 2023-03-24 src/tex/ms.tex
[showyourwork] overleaf sync

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. In particular, the past two decades revealed that planets can occur in orbits so close to their host star that they could be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions that emphasizes that they are bidirectional, affecting both planet, and star.

@@

\subsection{Magnetic star-planet interactions}

Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. It does not need to be there all the time, but for a time period that is longer than it takes for the effects of the planet to travel to the surface of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in some way in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby} and \citet{zarka2007plasma, saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all.

\subsection{Searching for global changes in actvity}

Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but just a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible,poppenhaeger2011correlation, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces a bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability.

\subsection{Flares}

In this study, we will look at flares as another way of dissipating the energy deposited in the magnetic field. Flares are explosions that occur in the stellar corona as a way to dissipate energy stored in the magnetic field by the motion of footpoints of coronal loops. They are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines~\citep{svestka1976solar,priest2002magnetic}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2} so that they can be detected in stars up to hundreds of parsecs away~(XXXX). But since they occur randomly in time, individual events can be hard to catch. %We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}.

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of these missions for this work is their quasi uninterrupted monitoring of star-planet systems that covers all phases of their planets' orbits, as will become clear in the following.

\subsection{Flaring star-planet interactions}

The absolute power of flaring star-planet interactions is difficult to constrain, so we do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might for instance be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism, so it might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

\subsection{Searching for local changes in activity}

Phase correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed a flare-up in X-ray when the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. Similarly, in \citet{ilin2022searching}, we found hints of excess flaring in the TESS light curves of the young AU Mic system. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}, and take a look at the more general patterns emerging from our systematic analysis~(Section~\ref{sec:discussion}). Finally, we summarize our findings, and suggest future observations that could uncover the star-planet interaction mechanisms operating behind the observed trends~(Section~\ref{sec:summary}).

@@

\section{Data}
\label{sec:data}
In this study, we wanted to compile the largest possible sample to search for flaring star-planet interactions, so took the entire Planetary Systems Composite Parameters Table (PSCP, Section~\ref{sec:data:sps})\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}} from the NASA Exoplanet Archive, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We ended up searching a total of 1812 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we then required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we also wanted to estimate the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation periods, planetary radii, and stellar luminosity~(Sections~\ref{sec:data:orbitalperiod}-\ref{sec:data:lum}). We started compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table from the literature. The literature parameters for the systems in the final sample of systems, for which we performed the full analysis, is shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeatable with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flare timing.

Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over 100000 stars monitored in 2-min cadence. 
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

The Transiting Exoplanet Survey Satellite~(TESS,~\cite{ricker2015transiting}) is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the 600-1000 nm band for $\sim 27 days$ in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations with about 20000 targets per Sector. Out of these, from Sector 27 on, 1000 targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we queried the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7258 light curves for flares -- 3030 Kepler Quarters, and 4228 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high costs of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

\subsection{Orbital periods and transit mid-times}
\label{sec:data:orbitalperiod}
We adopt the orbital periods from the PSCP, which were either obtained from Kepler or TESS transits, or from radial velocity measurements. For the transiting systems, we use the transit mid-times given in the PSCP to set orbital phase zero. If possible, we do this for each mission separately to reduce the uncertainties on orbital phase. For the non-transiting system, phase zero is set arbitrarily.
\subsection{Semi-major axes and eccentricities}
\label{sec:data:a}
The critical parameter is the distance between the star and the planet, not the semi-major axis itself. This varies if the orbit is eccentric. To account for varying distance, we adopt the mean semi-major axes from the literature, but use a custom estimate for the uncertainty that includes the eccentricity.

@@

GJ 3082 periods were measured and are consistent both in TESS and in KELT light curves~\citep{gunther2020stellar} to below 1\,s precision. GJ 674 periods are consistent between light curve periodograms~\citep{kiraga2007agerotationactivity} and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has measured rotation from activity measurements~\citep{astudillo-defru2017magnetic}, without any periodicities detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation period for Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

\subsection{Planetary radii}
\label{sec:data:planetradii}
For the transiting planets, we adopt the literature values for planet radius $R_p$, and uncertainties, from the PSCP. For the non-transiting radial velocity detected planets we use the planetary mass $M_p$ or $M_p\sin i$ to calculate the radius or a lower limit for the radius using the empirical relations derived in \cite{chen2017probabilistic} using their open source \texttt{forecaster} tool, upgraded to \texttt{astro-forecaster}\footnote{https://pypi.org/project/astro-forecaster/} by Ben Cassesse. We note that for GJ 674, the mass estimates in \cite{bonfils2007harps} and \cite{boisse2011disentangling} do not quote uncertainties, but differ by 0.3MEarth, so we assumed that value as the uncertainty on $M_p\sin i$.

@@

We take bolometric luminosity values as given in the PSCP whenever they are given with uncertainties. We supplement missing values, and replace entries without quoted uncertainties with Gaia DR3 FLAME~\citep{fouesneau2022gaia} solutions~(\texttt{lum\_flame}).

@@

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a p-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P_{SPI}$ in each system.

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. With relative velocity and magnetic field strength derived, we can combine them with stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws from two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process. First, we apply spline fit with a coarse sampling of 30h mean values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\cite{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for very active stars like AU Mic or Proxima Cen) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked.

In these light curves, we then search for flares as series of at least three consecutive data points $3\sigma$ above the noise level, implemented in AltaiPony~\cite{ilin2021altaipony}. We estimate the noise
level as the standard deviation in a rolling window of two hours, while masking deviation above $2.5\sigma$ (or $1.5$ sigma for active stars). To capture the exponential tail of the flare, we use the \texttt{addtail} flag to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold. For these flares, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{flare}(t)}{F_0}.
\end{equation}
Equivalent duration is the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in their Kepler and/or TESS light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}. In brief, we first take the number of flares observed in the TESS and Kepler data, calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been covered by the data, again for each light curve. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. This number of expected flares per bin can be aggregated into one expected distribution by summing over all available light curves. As a last step, this distribution is then compared to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Relative velocity}
\label{sec:methods:relvel}
We calculate the relative velocity between the magnetic field of the star, which we assume is co-rotating with the stellar rotation period $P_{rot}$, at the planetary orbit, where the planet moves with a period $P_{orb}$:

@@

\begin{equation}
    P_{sb} \sim B_p^{2/3} B_*^{4/3} v_{rel} R_p^2 a^{-4} R_*^4.
\end{equation}

$B_p$ and $B_*$ are the total planetary and stellar surface field strengths, respectively; $v_{rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis; and $R_*$ and $R_p$ are the stellar and planetary radii, respectively; and $a$ is the semi-major axis.

@@

\begin{equation}
    P_{sb0} \sim B_*^2 v_{rel}  R_p^2  a^{-4} R_*^4
\end{equation}

Third, for comparison, we estimate the power of star-planet interaction generated by the Alfv\'en wing mechanism ~\cite{saur2013magnetic,kavanagh2022radio}, combining Eqs.~8~and~11 in \cite{kavanagh2022radio} to give

\begin{equation}
    P_{aw} \sim R_{p}^2  B_p^{2/3}  B^{1/3}  v_{rel}^2 a^{-2}  
\end{equation}
where we approximate the stellar wind density $\rho_W$ with $a^{-2}$ and stellar wind magnetic field at the planetary radius $B_W$ with the surface field strength $B a^{-3}$, which result in the total $a^{-2}$ factor.

>>>>

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits so close to their host star that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Their effects pose challenges, but also opportunities for our understanding of star-planet systems:

@@

The search for magnetic SPI (Section~\ref{sec:intro:mspi}) in the form of changes in global activity indicators~(Section~\ref{sec:intro:local}), has so far been a series of detections and non-detections, often due to selection bias in the studied systems. Looking for local changes that trace the magnetic footpoint that connects planet to star mitigates this bias, but introduces another -- sampling bias due to poor orbital phase coverage. Flares~(Section~\ref{sec:intro:flares}) are activity markers that can be caused by SPI~(Section~\ref{sec:intro:fspi}), and be monitored more easily with high phase coverage than chromospheric lines or coronal brightness over long periods of time with missions like Kepler and TESS. Thanks to their enormous archives of hundreds of cumulative years of stellar monitoring, we can overcome the limitations of individual system studies~(Section~\ref{sec:intro:local}), and now look for flaring SPI more comprehensively, and systematically.

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in some way in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby} and \citet{zarka2007plasma, saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. All suffer from a lack of observational constraints for calibration.

\subsection{Searching for global changes in actvity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but just a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}.

\subsection{Flares}
\label{sec:intro:flares}
In this study, we will look at stellar flares as another way of dissipating the energy deposited in the magnetic field. Flares are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines in the corona~\citep{svestka1976solar,priest2002magnetic}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2} so that they can be detected in stars up to hundreds of parsecs away~(XXXX). But since they occur randomly in time, individual events can be hard to catch. %We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}.

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. As will become clear in the following Section, their key advantage for this work is the quasi uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

\subsection{Flaring star-planet interactions}
\label{sec:intro:fspi}
The absolute power of flaring star-planet interactions is difficult to constrain, so we do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might for instance be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism, so it might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
Phase correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed a flare-up in X-ray when the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. Similarly, in \citet{ilin2022searching}, we found hints of excess flaring in the TESS light curves of the young AU Mic system. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}). We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we wanted to compile the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP, Section~\ref{sec:data:sps})\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}} from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we then required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:orbitalperiod}-\ref{sec:data:lum}). We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table from the literature. The literature parameters for the systems in the final sample of systems, for which we performed the full analysis, is shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flare timing.

Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

The Transiting Exoplanet Survey Satellite~(TESS,~\cite{ricker2015transiting}) is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we use the \texttt{lightkurve}~\citep{lightkurvecollaboration2018lightkurve} software to query the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7258 light curves for flares -- 3030 Kepler Quarters, and 4228 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high costs of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

\subsection{Orbital periods and transit mid-times}
\label{sec:data:orbitalperiod}
We adopt the orbital periods from the PSCP, which were either obtained from Kepler or TESS transits, or from radial velocity measurements. For the transiting systems, we use the transit mid-times given in the PSCP to set orbital phase zero. If possible, we use the transit mid-times determined in each mission separately to reduce the uncertainties on orbital phase. For the non-transiting system, phase zero is set arbitrarily.
\subsection{Semi-major axes and eccentricities}
\label{sec:data:a}
The critical parameter is the distance between the star and the planet, not the semi-major axis itself. This varies if the orbit is eccentric. To account for varying distance, we adopt the mean semi-major axes from the literature, but use a custom estimate for the uncertainty that includes the eccentricity.

@@

The rotation of GJ 3082 was measured and are consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

\subsection{Planetary radii}
\label{sec:data:planetradii}
For the transiting planets, we adopt the literature values for planet radius $R_p$, and uncertainties, from the PSCP. For the non-transiting radial velocity detected planets we use the planetary mass $M_p$ or $M_p\sin i$ to calculate the radius or a lower limit for the radius using the empirical relations derived in \cite{chen2017probabilistic} using their open source \texttt{forecaster} tool, upgraded to \texttt{astro-forecaster}\footnote{https://pypi.org/project/astro-forecaster/} by Ben Cassesse. We note that for GJ 674, the mass estimates in \cite{bonfils2007harps} and \cite{boisse2011disentangling} do not quote uncertainties, but differ by $0.3M_\oplus$, so we assumed that value as the uncertainty on $M_p\sin i$.

@@

We take bolometric luminosity values as given in the PSCP whenever they are given with uncertainties. We supplement missing values, and replace entries without quoted uncertainties with Gaia DR3 FLAME~\citep{fouesneau2022gaia} solutions~(\texttt{lum\_flame, lum\_flame\_upper, lum\_flame\_lower}).

@@

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a p-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P$ in each system.

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws from two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for very active stars like AU Mic or Proxima Cen) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked.

In the de-trended light curves, we search for flares as series of at least three consecutive data points $3\sigma$ above the noise level, implemented in AltaiPony~\cite{ilin2021altaipony}. We estimate the noise level as the standard deviation in a rolling window of two hours, while masking deviation above $2.5\sigma$ (or $1.5$ sigma for active stars). To capture the exponential tail of the flare, we use the \texttt{addtail} flag in the \texttt{find\_flares} method to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold.

We vet all flare candidates by eye, and exclude instrumental false positives and physical one like Solar System Objects. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{flare}(t)}{F_0}.
\end{equation}
Equivalent duration is the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in their light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been covered by the data, also for each light curve separately. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare this distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Relative velocity}
\label{sec:methods:relvel}
We calculate the relative velocity between the magnetic field of the star and the planet using stellar rotation period $P_{rot}$, orbital period $P_{orb}$, and semi-major axis $a$. We assume that the large scale magnetic field is co-rotating with the stellar surface, and calculate the relative velocity at the planet's orbital distance:

@@

\begin{equation}
    P_{sb} \sim B_p^{2/3} R_p^2 B_*^{4/3} R_*^4 v_{rel} a^{-4} .
\end{equation}

$B_p$ and $B_*$ are the total planetary and stellar surface field strengths, respectively; $v_{rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.

@@

\begin{equation}
    P_{sb0} \sim R_p^2 R_*^4 B_*^2 v_{rel}    a^{-4} 
\end{equation}

Third, for comparison, we estimate the power of star-planet interaction generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, combining Eqs.~8~and~11 in \citet{kavanagh2022radio} to give

\begin{equation}
    P_{aw} \sim B_p^{2/3} R_{p}^2    B_*^{1/3}  v_{rel}^2 a^{-2}  
\end{equation}
where we approximate the stellar wind density $\rho_W$ with $a^{-2}$ and stellar wind magnetic field at the planetary radius $B_W$ with the surface field strength $B_* a^{-3}$, which result in the total $a^{-2}$ factor.
